
```
###Prerequisite
###Install the package if you haven't
setwd("~/git/")
require('devtools')
build('jkmeans')
install.packages("jkmeans_1.0.tar.gz", repos = NULL, type = "source")
```

```{r include=FALSE}
require('ggplot2')
require('reshape')
```



#Running jk-means, step-by-step

###1. Generate Data
Let's generate data from 3 clusters, each with n=100

N(1,1)
N(2,1)
N(6,1)

```{r}
mu<- rep(c(1,2,6),each=100)
table(mu)
y<- rnorm(100*3, mu, 1)
Z<- matrix(0,300,3)
clusterLabel<- rep(c(1:3),each=100)
for(i in 1:300){
  Z[i,clusterLabel[i]]<-1
}


hist(y,breaks=50)
```


###2. Run jk-means
Run with jk-means with j=2, k=3:

options: 
max iteration=100
convergence error tolerance = 1E-8
```{r}
K=3
j=2
jk <- jkmeans::jkmeansEM(as.matrix(y), K, j ,  1000, tol = 1E-8)

```

###3. Results
Here are the estimates for $\mu_k$ (sorted in ascending order)
```{r}
new_order<- order(jk$mu)
mu_est<- jk$mu[new_order]
```

Here are the cluster assignment probability: 

$$\zeta_{ik} = \frac{N(y_i|\mu_k,\sigma^2)}{\sum_k N(y_i|\mu_k,\sigma^2)}$$

```{r echo=FALSE}
#set up a coloring scheme using colorRampPalette
zeta_est<- jk$zeta[,new_order]
m <-melt(zeta_est)
names(m)<- c("i","k","prob")
p <- ggplot(data=m, aes(x=i, y=k, fill=prob)) + geom_tile()
p<- p+ scale_fill_continuous(low = "#FFFFFF", high="red", name = "")
p
```

###4. Validation


Compute RMSE
```{r}
RMSE <- function(a, b) {
  sqrt(mean((a - b) ^ 2))
}

RMSE(c(1,2,6),mu_est)

```

Use $\zeta_{ik}$ as  estimator for cluster assignment, compute the expected misclassification loss for each $\sum_k \zeta_{ik} (1- Z_{ik})$:

```{r}

#misclassified ones:
m <-rowSums( zeta_est*(1-Z))

#misclassification rate
sum(m)/300
```


Use $\hat{Z}_{i}= \max_k\zeta_{ik}$ as MAP estimator for cluster assignment, compute the misclassification error ($\sum_k \delta_{\hat{Z}_{i},k}(1-{Z_{ik}})$):

```{r}

misclsError<-function(zeta_est, n, Z){
  Z_est <- t(apply(zeta_est, MARGIN = 1, function(x){x==max(x)}))*1
  m <-rowSums(Z_est*(1-Z))
  sum(m)/n
}

#misclassification rate
misclsError(zeta_est,300, Z)
```


#Run jk-means with different n (repeat 50 times for each)

Define the functions:
```{r}



jkmeansExperiment<-function(n,j=3, sigma=1){
  
  
  mu<- rep(c(1,2,6),each=n)

  Z<- matrix(0,n*3,3)
  clusterLabel<- rep(c(1:3),each=n)
  for(i in 1:(n*3)){
    Z[i,clusterLabel[i]]<-1
  }

  y<- as.matrix(rnorm(n*3, mu, sigma))
  
  jk <- jkmeans::jkmeansEM(y, K, j=j ,  1000, tol = 1E-8)
  new_order<- order(jk$mu)
  mu_est<- jk$mu[new_order]
  zeta_est<- jk$zeta[,new_order]
c(RMSE(c(1,2,6),mu_est),misclsError(zeta_est,3*n,Z))
}

runExp<- function(j=3,n=100, m=50){
  
  
  rmse<- numeric(m)
  mce<- numeric(m)

  for(i in 1:m){
      result<- jkmeansExperiment(n,j=j)
      rmse[i]<- result[1]
      mce[i]<- result[2]
  }
  return(list("RMSE"=rmse,"MCE"=mce))
}

```


Run n=seq(10,500,by=10), get average RMSE and MCE over 50 times for each n:

run with $j=3$
```{r}

n_seq<- seq(10,500,by=10)


avg_rmse_3<- numeric(length(n_seq))
avg_mce_3<- numeric(length(n_seq))

for(i in 1:length(n_seq)){
  n<- n_seq[i]
  Exp<- runExp(j=3,n=n,m=50)
  avg_rmse_3[i]<- mean(Exp$RMSE)
  avg_mce_3[i]<- mean(Exp$MCE)
  print(i)
}

```

run with $j=2$
```{r}


avg_rmse_2<- numeric(length(n_seq))
avg_mce_2<- numeric(length(n_seq))

for(i in 1:length(n_seq)){
  n<- n_seq[i]
  Exp<- runExp(j=2,n=n,m=50)
  avg_rmse_2[i]<- mean(Exp$RMSE)
  avg_mce_2[i]<- mean(Exp$MCE)
  print(i)
}

```


run with $j=1$
```{r}


avg_rmse_1<- numeric(length(n_seq))
avg_mce_1<- numeric(length(n_seq))

for(i in 1:length(n_seq)){
  n<- n_seq[i]
  Exp<- runExp(j=1,n=n,m=50)
  avg_rmse_1[i]<- mean(Exp$RMSE)
  avg_mce_1[i]<- mean(Exp$MCE)
  print(i)
}

```



Plot of the RMSE vs n

```{r}

plot1<- data.frame( "n" = rep(n_seq, 3),
            "RMSE_avg"=c(avg_rmse_1,avg_rmse_2,avg_rmse_3),
            "j"= as.factor(rep(c(1:3),each= length(n_seq)))
            )

p<- ggplot(data=plot1, aes())
p<- p+  geom_line(aes(x= n, y=RMSE_avg, colour=j)) 
p<- p+ theme_bw() + labs(title = "mu = 1,2,6")
p
```


Plot of the misclassification error vs n

```{r}

plot2<- data.frame( "n" = rep(n_seq, 3),
            "MCE_avg"=c(avg_mce_1,avg_mce_2,avg_mce_3),
            "j"= as.factor(rep(c(1:3),each= length(n_seq)))
            )

p<- ggplot(data=plot2, aes())
p<- p+  geom_line(aes(x= n, y=MCE_avg, colour=j)) 
p<- p+ theme_bw() + labs(title = "mu = 1,2,6")
p
```

